### 2012-09-08 bidir_pathtrace_skeleton.tex
Quelques corrections supplémentaires

\documentclass[final,envcountsame]{llncs}

%\usepackage{graphicx} %Pouvoir insérer des images
\usepackage{epsfig}
\usepackage{amsmath} %Permet d'utiliser une jolie panoplie de symboles mathématiques
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsfonts}

\def\axisloc#1{\mathcal{LOC}_{#1}}
\def\axislocbis#1#2{\mathcal{LOC}_{#1}^{>#2}}
\def\axislc#1{\mathcal{LC}_{#1}}
\newcommand{\centerness}{Decenter}

\def\OneBall#1{\mathbb{B}^{1}_{#1}}
\def\MaxOneBall#1{\mathbb{M}\OneBall{#1}}

\def\dd{\omega}
\def\DD{\Omega}

\def\mydef#1{{\em #1}}
\def\myem#1{{\em #1}}
\def\myvec#1{\vec{#1}}
\def\quotes#1{``#1''}

\def\birth#1#2{Birth_{#2}(#1)}
\def\death#1#2{Death_{#2}(#1)}
\def\lifespan#1#2{Lifespan_{#2}(#1)}

\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\cont#1{\check{#1}}
\def\contstrict#1{\check{#1}^*}
\def\closure#1{#1^-}
\def\princ#1{#1^+}
\def\detach#1#2{#2 \oslash #1}
\def\attach#1#2{\mbox{\it{}Attach\/}(\cell{#1},#2)}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}Border\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{\overline{#1}}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

\def\figpath#1{../../Images/EPS/#1}

\begin{document}

\mainmatter

\title{A 3d curvilinear skeletonization algorithm with application to bidirectional path tracing}
\author{John Chaussard \and Venceslas Biri \and Michel Couprie}
\institute{Universit\'{e} Paris Est, LABINFO-IGM, A2SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{chaussaj@esiee.fr, bertrang@esiee.fr, coupriem@esiee.fr}}

\maketitle

\begin{abstract}
What we propose is really great...
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                        %
%                   INTRODUCTION                         %
%                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec::intro}
Blah blah blah...

\section{The cubical complex framework}

\subsection{Basic definitions}

In the 3d voxel framework, objects are mae of voxels. In the 3d cubical complex framework, objects made of voxel, but of cubes, squares, lines and vertices. 
Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$. Any subset $f$ of $\Z^n$ such that $f$ is the cartesian product of $m$ elements of $\mathbb{F}^1_1$ and $(n-m)$ elements of $\mathbb{F}^1_0$ is called a face or an \myem{$m$-face} of $\Z^n$, $m$ is the dimension of $f$, we write $dim(f)=m$. A $0$-face is called a \myem{vertex}, a $1$-face is an \myem{edge}, a $2$-face is a \myem{square}, and a $3$-face is a \myem{cube}.

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.

Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ is \myem{a face of $f$}, and any element of $\cellstrict{f}$ is \myem{a proper face of $f$}. We call \myem{star of $f$} the set $\cont{f} = \{g \in \allfaces{n} \vert f \subseteq g\}$, and we write $\contstrict{f} = \cont{f} \setminus \{f\}$: any element of $\cont{f}$ is a \myem{coface of $f$}. It is plain that $g\in\cell{f}$ iff $f\in\cont{g}$.

A set $X$ of faces in $\allfaces{n}$ is a \myem{cell}, or \myem{$m$-cell}, if there exists an $m$-face $f \in X$ such that $X=\cell{f}$. The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$. The set $\Compl{X}$ is $\allfaces{n} \setminus X$.

\begin{definition}
A finite set $X$ of faces in $\allfaces{n}$ is a \myem{cubical complex} if $X=\closure{X}$, and we write $\complex{n}{X}$.

Any subset $Y$ of $X$ which is also a complex is a \myem{subcomplex of $X$}, and we write $\subcomplex{Y}{X}$.
\end{definition}

Informally, in 3d, a set of faces $X$ is each side (square) of a cube belonging to $X$ also belong to $X$, each edge of a square belonging to $X$ also belong to $X$, and each vertex of an edge belonging to $X$ also belong to $X$.

A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$. We denote by $\princ{X}$ the set composed of all facets of $X$ . A complex $X$ is \myem{pure} if all its facets have the same dimension. The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$. If $\mydim{X} = d$, then we say that $X$ is a $d$-complex.
In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$; in $\allfaces{3}$, a complex is thin if it contains no cube.

\subsection{From binary images to cubical complex}
\label{sec::binary_to_cubic}

Traditionally, a voxel image is defined as a finite subset of $\Zset^n$. This kind of image is the most common one in the field of image processing and we give now a simple way to transpose a binary image to the cubical complex framework.

Informally, to do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$ (to a pixel we associate a square, to a voxel we associate a cube). More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\Phi(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$. We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$. Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$.

\subsection{Thinning: the collapse operation}
The collapse operation is the basic operation for performing homotopic thinning of a complex. It consists of removing two distinct elements $(f,g)$ from a complex $X$ under the condition that $g$ is contained in $f$ and is not contained in any other element of $X$. This operation may be repeated several times.

\begin{definition}
Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$. The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $f$ is the only face of $X$ such that $g$ is a proper face of $f$. 
\end{definition}

In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$. It can be easily seen that if $(f,g)$ is a free pair for a complex $X$ and $\dim(f)=m$, then $f$ is a facet and $\dim(g) = m-1$.

\begin{definition}
Let $\complex{n}{X}$, and let $(f,g)$ be a free pair for $X$. The complex $X \setminus \{f,g\}$ is an \myem{elementary collapse of $X$}.

Let $\complex{n}{Y}$, the complex $X$ \myem{collapses onto} $Y$ if there exists a sequence of complexes $(X_0,...,X_\ell)$ of $\allfaces{n}$ such that $X=X_0$, $Y=X_\ell$ and for all $i \in \dinterval{1}{\ell}, X_{i}$ is an elementary collapse of $X_{i-1}$. We also say, in this case, that $Y$ \myem{is a collapse of} $X$.
\end{definition}

If $Y$ is a collapse of a complex $X$, then $Y$ is a complex. 
We now introduce some elements that will serve for proving the thinness of our skeletons.
Let $f_0,f_\ell$ be two $n$-faces of $\allfaces{n}$ (with $\ell$ being even). An \myem{$(n-1)$-path from $f_0$ to $f_\ell$} is a sequence $\pi =
(f_0,...,f_\ell)$ of faces of $\allfaces{n}$ such that for all $i \in \dinterval{0}{\ell}$, either $i$ is even and $f_i$ is an $n$-face, or $i$ is odd and $f_i$ is an $(n-1)$-face with $\contstrict{f_i}=\{f_{i-1}, f_{i+1}\}$ (such path always exists).

\begin{proposition}
\label{prop:thinness}
Let $\complex{n}{X}$ be an $n$-complex, with $n > 0$. Then $X$ has at least one free $(n-1)$-face.
\end{proposition}

\begin{proof}
Since $X$ is an $n$-complex (hence $X$ is finite) there exists an $n$-face $a$ in $X$ and an $n$-face $b$ in $\Compl{X}$. Obviously, there exists an $(n-1)$-path from $a$ to $b$. Let $h$ be the first $n$-face of $\pi$ that is not in~$X$, let $k$ be the last $n$-face of $\pi$ before $h$ (thus $k$ is in $X$), and let $e = k \inter h$ be the $(n-1)$-face of $\pi$ between $k$ and $h$. Since $k$ and $h$ are the only two $n$-faces of $\allfaces{n}$ that contain $e$, we see that the pair $(k,e)$ is free for $X$. $\finproof$
\end{proof}

In conclusion, in $\allfaces{3}$, as long as a complex still contains $3$-faces, it has a free $2$-face and more collapse operations can be performed. Therefore, it is possible to perform collapse on a complex until no more $3$-faces (volumes) can be found (until it is thin).

\section{A parallel directional thinning based on cubical complex}

\subsection{Removing free pairs in parallel}
In the cubical complex framework, parallel removal of simple pairs can be easily achieved when following simple rules that we give now. First, we need to define the \myem{direction} and the \myem{orientation} of a free face.

Let  $f \in \allfaces{n}$, the \myem{center of $f$} is the center of mass of the points in $f$, that is, $c_f= \frac{1}{\Card{f}}\sum_{a \in f}{a}$. The center of $f$ is an element of $[\Zhalf]^n$, where $\Zhalf$ denotes the set of half integers.
Let $\complex{n}{X}$, let $(f,g)$ be a free pair for $X$, and let $c_f$ and $c_g$ be the respective centers of the faces $f$ and $g$. We denote by $V(f,g)$ the vector $(c_f -c_g)$ of $[\Zhalf]^n$.

We define a surjective function $\type{}: \allfaces{n} \times \allfaces{n} \rightarrow \dinterval{0}{n-1}$ such that, for all free pairs $(f,g)$ and $(i,j)$ for $X$, $\type{f,g} = \type{i,j}$ if and only if $V(f,g)$ and $V(i,j)$ are collinear (we don't bother defining $\type{}$ for non free pairs as it won't be useful in this case). The number $\type{f,g}$ is called the \myem{direction} of the free pair $(f,g)$. Let $(f,g)$ be a free pair, the vector $V(f,g)$ has only one non-null coordinate: the pair $(f,g)$ has a \myem{positive orientation}, and we write $\orient{f,g}=1$, if the non-null coordinate of $V(f,g)$ is positive; otherwise $(f,g)$ has a \myem{negative orientation}, and we write $\orient{f,g}=0$.
%On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but opposite orientations.

Now, we give a property of collapse which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology.

\begin{proposition}
\label{prop:collpar2}
Let $\complex{n}{X}$, and let $(f,g)$ and $(k,\ell)$ be two distinct free pairs for $X$. The complex $X$ collapses onto $X \setminus \{f,g,k,\ell\}$ if and only if $f \neq k$. 
\end{proposition}

\begin{proof}
If $f=k$, then it is plain that $(k,\ell)$ is not a free pair for $Y=X \setminus \{f,g\}$ as $k=f \notin Y$. Also, $(f,g)$ is not free for $X \setminus \{k,\ell\}$. If $f\neq k$, then we have $g \neq \ell$, $\contstrict{g} \inter X = \{f\}$ ($g$ is free for $X$) and $\contstrict{\ell} \inter X = \{k\}$ ($\ell$ is free for $X$). Thus, we have $\contstrict{\ell} \inter Y = \{k\}$ as $\ell\neq g$ and $k\neq f$. Therefore, $(k,\ell)$ is a free pair for $Y$. $\finproof$
\end{proof}

From Prop.~\ref{prop:collpar2}, the following corollary is immediate.

\begin{corollary}
\label{cor:parallel_removal}
Let $\complex{n}{X}$, and let $(f_1,g_1) \ldots (f_m,g_m)$ be $m$ distinct free pairs for $X$ such that, for all $a,b \in
\dinterval{1}{m}$ (with $a \neq b$), $f_a \neq f_b$. The complex $X$ collapses onto $X \setminus \{f_1,g_1 \ldots f_m,g_m\}$.
\end{corollary}

Considering two distinct free pairs $(f,g)$ and $(i,j)$ for $\complex{n}{X}$ such that $\type{f,g}=\type{i,j}$ and $\orient{f,g} = \orient{i,j}$, we have $f \neq i$.  From this observation and Cor.~\ref{cor:parallel_removal}, we deduce the following property.

\begin{corollary}
\label{cor:parallel_removal_v2}
Let $\complex{n}{X}$, and let $(f_1,g_1) \ldots (f_m,g_m)$ be $m$ distinct free pairs for $X$ having all the same direction and the same orientation. The complex $X$ collapses onto $X \setminus \{f_1,g_1 \ldots f_m,g_m\}$. 
\end{corollary}

\subsection{A directional parallel thinning algorithm}

Intuitively, we want the algorithm to remove free faces of a complex \quotes{layer by layer}: we don't want to have unequal thinning of the input complex. Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex. 
We say that a $d$-face of $X$ is a \myem{border face} if it contains a free $(d-1)$-face. We define $\border{X}$ as the set of all border faces of $X$. We now  introduce a directional parallel thinning algorithm (Alg.~\ref{algo::pardircollapse_v0}).

On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}). Thanks to corollary \ref{cor:parallel_removal_v2}, we also remove faces with same direction and orientation in parallel (l.~\ref{pc::08}).

\begin{algorithm}[h!]
\label{algo::pardircollapse_v0}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[ParDirCollapse]{$ParDirCollapse(X,W,\ell)$}
\KwData{A cubical complex $\complex{n}{X}$, a subcomplex
  $\subcomplex{W}{X}$ which represents faces of $X$ which should not
  be removed, and $\ell \in \Nset$, the number of layers of free faces
  which should be removed from $X$} 
\KwResult{A cubical complex}
\lnl{pc:01}\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{
\lnl{pc:02}  $L=\border{X}^-$\;
\lnl{pc:03}  \For{$t=0 \rightarrow n-1$}{
\lnl{pc:04}    \For{$s=0 \rightarrow 1$}{
\lnl{pc:05}      \For{$d=n \rightarrow 1$}{
\lnl{pc:06}        $E=\{(f,g)$ free for $X \suchthat g \notin W,$\\
                      $\;$~~~~~~~$\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;
\lnl{pc:07}        $G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\;
\lnl{pc:08}        $X=X \setminus G$\;
                 }
               }
             }
\lnl{pc:09}  $l = l-1$\;
	   }
\lnl{pc:10} \Return $X$\;
\end{algorithm}

Different definitions of the orientation and direction can be given, corresponding to different order of removal of free faces in the complex and leading to different results.
Algorithm~\ref{algo::pardircollapse_v0} may be easily implemented to run in linear time complexity (proportionally to the number of faces of the complex). Indeed, checking if a face is free or not may be easily done in constant time. Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the elements of $\cellstrict{f} \union \contstrict{g}$ in order to find new free faces, as other faces' status won't change.

The following property is a direct consequence of of Prop.~\ref{prop:thinness} and the fact that Alg.~\ref{algo::pardircollapse_v0} stops when no more free faces can be found in the complex :

\begin{proposition}
\label{prop::alg_v0_thin}
Let $\complex{n}{X}$, the result of $ParDirCollapse(X,\emptyset,+\infty)$ is a thin complex.
\end{proposition}

\section{Aspect preservation during thinning: a parameter-free method}
\label{sec::aspect_no_param}
Algorithm \ref{algo::pardircollapse_v0} does not necessarily preserves "geometrical information" from the original object in the resulting skeleton. For example, if the input was a filled human shape (one connected component, no tunnel nor cavity) and that $W=\emptyset$ and $l=+\inf$, then the result of Alg.~\ref{algo::pardircollapse_v0} would be a simple vertex. However, when one wants to preserve "geometrical information" from the original shape in the skeleton, then one expects to obtain a "stickman" from a human shape. However, it is important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.

In the following, we will see a new method in the cubical complex, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory visual properties. This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0}, some faces from the original object during the thinning process.

\subsection{The lifespan of a face}
In the following, we define more notions in the cubical complex. The first one we present is the \myem{death date} of a face.

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{death date of $f$ in $X$}, denoted by $\death{f}{X}$, is the smallest integer $d$ such that $f \notin ParDirCollapse(X, \emptyset, d)$. 
\end{definition}

The death date of a face indicates how many layers of free faces should be removed from a complex $X$, using Alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$. We now define the \myem{birth date} of a face:

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or either $f \notin ParDirCollapse(X, \emptyset, b)$.
\end{definition}

The birth date indicates how many layers of free faces must be removed from $X$ with Alg.\ref{algo::pardircollapse_v0} before transforming $f$ into a facet of $X$ (we consider a face "lives" when it is a facet).
Finally, we can define the \myem{lifespan} of a face :

\begin{definition}
Let $f \in \complex{n}{X}$, the lifespan of $f$ in $X$ is the integer

\begin{math}
\lifespan{f}{X} = \left\{
\begin{array}{ll}
+\infty & \text{if $\death{f}{X}=+\infty$}\\
\death{f}{X} - \birth{f}{X} & \text{otherwise}
\end{array}
\right.
\end{math}
\end{definition}

The three parameters previously defined are dependant on the order of direction and orientation chosen for algorithm $ParDirCollapse$.

The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with Alg.~\ref{algo::pardircollapse_v0}.

The lifespan is a good indicator of how important a face can be in an object. Typically, higher the lifespan is, and more representative of an object's visual feature the face is. The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.

\subsection{Distance map and opening function}
\label{subsec::distmap}
In addition to the lifespan of a face, the proposed homotopic thinning method uses information on distance between faces in order to decide if a face should be kept safe from deletion. We define hereafter the various notions needed for this, based on distance in the voxel framework.

Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal or inferior to $1$ (also called direct neighbours). A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours. The length of $\mathcal{C}$ is $k$. Remark that $d_1$ is indeed the $6$-distance in the 3d voxel framework.

We set $d_1(x,y)$ as the length of the shortest $1$-path from $x$ to $y$. Let $S \subset \Zset^n$, we set $d_1(x,S) = \displaystyle \min_{y \in S} d_1(x,y)$. The \myem{$1$-ball of radius $r$ centered on $x$} is the set $\OneBall{r}(x)=\{y \in \Zset^n | d_1(x,y) < r\}$. 
Given $X \subset \Zset^n$, the \myem{maximal $1$-ball of $X$ centered on $x$} is the set $\MaxOneBall{X}(x) = \OneBall{d_1(x,\Compl{X})}(x)$.

We set, for all $x \in X$, $\dd_1(x,\Compl{X}) = \displaystyle \max_{x \in \MaxOneBall{X}(y)} d_1(y,\Compl{X})$: this value indicates the radius of the largest maximal $1$-ball contained in $X$ and containing $x$. If $x \in \Compl{X}$, we set $\dd_1(x,\Compl{X}) = 0$. The map $\dd_1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.

Given $X \subset \Zset^n$, the value of $\dd_1(x,\Compl{X})$ of every $x \in X$ can be naively computed by performing successive morphological dilations of values of the map $d_1$. 
%TODO Parler du fait que l'on a une version dans la thèse, et une version plus amélioré dans thèse et public avenir

In order to extend $d_1$ and $\dd_1$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi : \subfaces{n}{n} \rightarrow \Zset^{n}$ defined in Sec.~\ref{sec::binary_to_cubic}. It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$. We indifferently use $\Phi^{-1}$ as a map from $\subfaces{n}{n}$ to $\Zset^n$, and as a map from $\mathcal{P}(\subfaces{n}{n})$ to $\mathcal{P}(\Zset^n)$.

Given $Y \subset \allfaces{n}$, we define the map $\tilde{D_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $d_1$ to the cubical complex framework: for all $f \in \allfaces{n}$,

\bigskip
\begin{math}
\tilde{D_1}(Y)(f) = \left\{
\begin{array}{ll}
d_1(\Phi^{-1}(f),\Phi^{-1}(Y \inter \subfaces{n}{n})) & \text{if $f$ is an $n$-face}\\
\displaystyle \max_{g \in \contstrict{f} \inter \subfaces{n}{n}} \tilde{D_1}(Y)(g) & \text{else}
\end{array}
\right.
\end{math}

Informally, if $f$ is a 3-face, then $\tilde{D_1}(Y)(f)$ is the length of the shortest 1-path between the voxel \quotes{corresponding} to $f$ and the set of voxels corresponding to $Y$.
The same way, we define $\tilde{\DD_1}(Y): \allfaces{n} \rightarrow \Nset$ as an extension of $\dd_1$ to the cubical complex framework.

Finally, we define the decenterness map:

\begin{definition}
The \myem{decenterness map}, denoted by $\centerness(Y)$, is equal to $\tilde{\DD_1}(Y) - \tilde{D_1}(Y)$.
\end{definition}

\subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
As previously said, we will add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object. If an edge has a high decenterness value, then it is probably located too close to the object's border and does not represent an interesting visual feature to preserve.
%TODO ref vers image de decenterness
On the other hand, if an edge has a high lifespan, then it means it was not removed quickly, after becoming a cell, by the thinning algorithm and might represent some precious visual information on the original object. An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value. Unfortunately, this strategy produces skeletons with many spurious branches in surfacic areas of the original object.

We can identify surfacic areas of a complex as zones where squares have a high lifespan. Therefore, in order to avoid spurious branches in surfacic areas, we need to make it harder for edges to appear in these zones. It can be achieved by deciding that an edge will be kept safe from deletion by the thinning algorithm if its lifespan is superior to the decenterness value plus the lifespan of squares \quotes{around} this edge. This leads us to

\begin{algorithm}[ht]
\label{algo::3d_input_curv_skel}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[CurvilinearSkeleton]{$CurvilinearSkeleton(X)$}
\KwData{A cubical complex $\complex{3}{X}$}
\KwResult{A cubical complex $\complex{3}{Y}$ such that $\dim{Y} \leq 2$}
$W=\axisloc{1}(X)$\;
$ParDirCollapse(X,W,+\infty)$\;
\Return $X$\;
\end{algorithm}

\input{pathtracint_application_skeleton}

\end{document}